API Evolution: A Study of Software Releases and Change Trends Leading to Stability

### 2013-11-18 apie-2013/paper.tex
APIE added Alessias changes. DRAFT 2 done.

While measuring software quality has had a major focus in software engineering research for many years~\cite{Bowen:1978:CAS}~\cite{Grady:1993:PRM}~\cite{ISOIEC9126},
the sub study of software stability and its implications on quality and reliability remains a difficult subject to understand. 
The decision of what makes a project stable and ready
for a release often comes down to the release manager or maintainer of a project and is often a reflection of the open source community which surrounds 
the project~\cite{Conway:1968}, or a reduction in code churn. Code is an often looked to statistic for stability but can be grossly misleading
in terms of pre-release and post-release defects~\cite{Fenton:2000:QAF}, with some exceptions~\cite{Nagappan:2005:URC}. 
Creating an approach to determining software stability and release preparedness is still a large open area of
interest in software engineering research.

@@

The remainder of this paper is laid out as follows. Section~\ref{sec:rel} goes over the related work of release study and stability. Sections~\ref{sec:meth}
explains our methodology and tools used in solving our research problem.
Section \ref{sec:results} gives a full representation of the results and trends found as well as a summarized results of the 9 change trends found
to be most prominent. Finally, Sections~\ref{sec:fut} and~\ref{sec:con} give our future work to follow this study and final conclusions respectively.

\section{Related Work}
\label{sec:rel}
While very little has been published about release quality studies and stability, there have been a few studies which attempt to address the issues directly
or indirectly. Wu et al.~\cite{Wu:2008:QAF} performed a case study of SoftPM, a widely adopted project management tool, to explore the relationships of
pre-release and post-release failures at major releases. Wu et al. found that the ratio of post-release failures to pre-release failures is significantly low
and can be used to show reliability and stability. Hindle et al.~\cite{Hindle:2007:RPD} performed a case study on MySQL which observed a the project's behavior
around major and minor release by monitoring artifact check-ins and changes. They found that there are temporary stoppages for source revisions around releases,
indicating that a temporary freeze is taking place for developers and that last minute fixes and manual testing may be being performed.
Zaidman et al.~\cite{Zaidman:2011:SCP}, in comparison, studied the co-evolution of production and test code with inspections and analysis
at major and minor releases.

The study of open source projects revolving around release points has become more available by the work of Tsay et al~\cite{Tsay:2011:EMO}. Tsay et al. created
a resource of historical release dates for open source software projects to be used for future studies by other researchers.

In terms of software stability, many proposed development techniques have been proposed to increase software stability. 
Fayad~\cite{Fayad:2001:TOI}~\cite{Fayad:2002:ASS} suggests that ``business objects'' (BOs) do not change in nature and that they are inherently stable. These
objects only need to change to accommodate external modules at the interface. Some studies such as Chow et al.~\cite{Chow:2011:SJI} have investigated
the stability of changes to interfaces which are considered a good indication of stability.
Mockus et al.~\cite{Mockus:2008:IQR} used major and minor release points to compare industry process quality to customer-perceived quality of the software
project. Mockus et al. found that defect density is lowest at major releases but at the same time software quality is at its lowest all when compared to minor
releases. The low software quality here related to end-user errors of installations and configurations. Wermelinger et al.~\cite{Wermelinger:2008:AEE}
showed that stable core architectures can be detected by using source code changes. Finally, Fayad et al.~\cite{Fayad:2010:SSM} have
investigated the Software Stability Model (SSM) for Software Product Lines to show that the SSM's impact on architecture and design of a software product
can help improve the life of the product line and make it more adaptable and applicable.

@@

Once the ChangeDistiller information was collected, we decided to examine software change trends surrounding releases of the project's we had selected. Since releases
have preconceived notions of software stability, we decided that by studying the types of changes surrounding these releases, we could get a better understanding of
what types of source code changes or trends constitute software stability or maturity. In order to study the release points, we went to each of our 10 project's 
web pages and looked through their release histories for major, minor, alpha, beta, and release candidate type releases. In total we identified 472 releases
across our 10 studied projects.

@@

Equation~\ref{eq:norm} works to create a change ratio by first creating a numerator by summing across all releases of
a given release type a sum of a particular change type in commits ($T_c$)
from the release date ($r$) to a given number of days after the release ($d$) divided by the number of commits in this date range ($|c|$). Next the denominator
is created by summing across all release of a given release type
a sum of that same particular change type in commits($T_c$) from a given number of days ($d$) before the release date ($r$) to
the release date divided by divided by the number of commits in this date range ($|c|$). This numerator and denominator form the final change ratio.
This equation gives us a ratio of a particular change type happening before and after a particular release. If the ratio is above 1 then that particular change
type occurs more frequently after the release and if it is below 1 then is occurs more frequently before the release. For the purposes of our study, we set the
number of days before and after the release ($d$) to 60 as the projects studies had many months in between their major releases. This quantitative data
formed much of the basis for the results to come in Section~\ref{sec:results}

@@

% APIE graph figure
\begin{figure}[tb!]
\centering
\includegraphics[width=0.5\textwidth]{images/apie.png}
\caption{An screen shot of the APIE visualizer showing project Eclipse.Mylyn.Context with change type PUBLIC\_ADDED\_METHODS being analyzed and showing
major releases as vertical yellow lines.\label{fig:apie}}
\end{figure}

@@

Due to the space requirements of this paper, we focus our results on Major releases of the 10 case study projects and a select few of the calculated change ratios. 
There were 109 major releases across the 10 studied projects. All of the major findings as per values computed from Equation~\ref{eq:norm} for non test metrics
can be seen in Table~\ref{tab:ratio}.

@@

As it can be seen in Table~\ref{tab:ratio}, there a few change type trends around major releases which pass the threshold. We can see that both public
and private methods
being removed from a project is more likely to occur after a major release than after. Table~\ref{tab:ratio} also shows significance in the changes to private
classes. We see that private classes are added more (24\%) before a major release and removed more after (44\%) the release. 
All results in Table~\ref{tab:ratio} could be used as identified trends of major software releases, while we have just highlighted the larger ratios
which meet our threshold criteria in this paper.

>>>>

While measuring software quality has had a major focus in software engineering research for many years~\cite{Bowen:1978:CAS}~\cite{Grady:1993:PRM}~\cite{ISOIEC9126},
the sub study of software stability and its implications on quality and reliability remains a difficult subject to understand. 
The decision of what makes a project stable and ready
for a release often comes down to the release manager or maintainer of a project and is often a reflection of the open source community which surrounds 
the project~\cite{Conway:1968}. Code churn is an often looked to statistically for stability but can be grossly misleading
in terms of pre-release and post-release defects~\cite{Fenton:2000:QAF}, with some exceptions~\cite{Nagappan:2005:URC}. 
Creating an approach to determining software stability and release preparedness is still a large open area of
interest in software engineering research.

@@

The remainder of this paper is laid out as follows. Section~\ref{sec:rel} goes over the related work of release study and stability. Sections~\ref{sec:meth}
explains our methodology and tools used in solving our research problem.
Section \ref{sec:results} gives a full representation of the results and trends found as well as the 9 change trends found
to be most prominent. Finally, Sections~\ref{sec:fut} and~\ref{sec:con} give our future work to follow this study and final conclusions respectively.

\section{Related Work}
\label{sec:rel}
While very little has been published about release quality studies and stability, there have been a few studies which attempt to address the issues directly
or indirectly. Wu et al.~\cite{Wu:2008:QAF} performed a case study of SoftPM, a widely adopted project management tool, to explore the relationships of
pre-release and post-release failures at major releases. Wu et al. found that the ratio of post-release failures to pre-release failures is significantly low
and can be used to show reliability and stability. Hindle et al.~\cite{Hindle:2007:RPD} performed a case study on MySQL which observed a project's behavior
around major and minor release by monitoring artifact check-ins and changes. They found that there are temporary stoppages for source revisions around releases,
indicating that a temporary freeze is taking place for developers and that last minute fixes and manual testing may be performed.
Zaidman et al.~\cite{Zaidman:2011:SCP}, in comparison, studied the co-evolution of production and test code with inspections and analysis
at major and minor releases, showing how test and production code can evolve at different rates and times. These results show that production
and test code should be handled as different cases for a stability measure around major releases.

The study of open source projects revolving around release points has become more accessible by the work of Tsay et al~\cite{Tsay:2011:EMO}. Tsay et al. created
a resource of historical release dates for open source software projects to be used for future studies by other researchers.

In terms of software stability, development techniques have been proposed to increase software stability. 
Fayad~\cite{Fayad:2001:TOI}~\cite{Fayad:2002:ASS} suggests that ``business objects'' (BOs) do not change in nature and that they are inherently stable. These
objects only need to change to accommodate external modules at the interface. Some studies such as Chow et al.~\cite{Chow:2011:SJI} have investigated
the stability of changes to interfaces which are considered a good indication of stability.
Mockus et al.~\cite{Mockus:2008:IQR} used major and minor release points to compare industry process quality to customer-perceived quality of the software
project. Mockus et al. found that defect density is lowest at major releases but at the same time software quality is at its lowest all when compared to minor
releases. The low software quality here relates to end-user errors of installations and configurations. Wermelinger et al.~\cite{Wermelinger:2008:AEE}
showed that stable core architectures can be detected by using source code changes. Finally, Fayad et al.~\cite{Fayad:2010:SSM} have
investigated the Software Stability Model (SSM) for Software Product Lines to show that the SSM's impact on architecture and design of a software product
can help improve the life of the product line and make it more adaptable and applicable.

@@

Once the ChangeDistiller information was collected, we decided to examine software change trends surrounding releases of the project's we had selected. Since releases
have preconceived notions of software stability, we decided that by studying the change types surrounding these releases, we could get a better understanding of
what types of source code changes or trends constitute software stability or maturity. In order to study the release points, we went to each of our 10 project's 
web pages and looked through their release histories for major, minor, alpha, beta, and release candidate type releases. In total we identified 472 releases
across our 10 studied projects.

@@

Equation~\ref{eq:norm} works to create a change ratio by first creating a numerator by summing across all releases of
a given release type a sum of a particular change type in commits ($T_c$)
from the release date ($r$) to a given number of days after the release ($d$) divided by the number of commits in this date range ($|c|$). Next the denominator
is created by summing across all release of a given release type
a sum of that same particular change type in commits ($T_c$) from a given number of days ($d$) before the release date ($r$) to
the release date divided by the number of commits in this date range ($|c|$). This numerator and denominator form the final change ratio.
This equation gives us a ratio of a particular change type happening before and after a particular release. If the ratio is above 1 then that particular change
type occurs more frequently after the release and if it is below 1 then it occurs more frequently before the release. For the purposes of our study, we set the
number of days before and after the release ($d$) to 60 as the projects studies had many months in between their major releases. This quantitative data
formed much of the basis for the results to come in Section~\ref{sec:results}

@@

% APIE graph figure
\begin{figure}[tb!]
\centering
\includegraphics[width=0.5\textwidth]{images/apie.png}
\caption{A screen shot of the APIE visualizer showing project Eclipse.Mylyn.Context with change type PUBLIC\_ADDED\_METHODS being analyzed and showing
major releases as vertical yellow lines.\label{fig:apie}}
\end{figure}

@@

Due to space requirements of this paper, we focus our results on major releases of the 10 case study projects and select few of the calculated change ratios. 
There were 109 major releases across the 10 studied projects. All of the major findings as per values computed from Equation~\ref{eq:norm} for non test metrics
can be seen in Table~\ref{tab:ratio}.

@@

As it can be seen in Table~\ref{tab:ratio}, there are few change type trends around major releases which pass our threshold. We can see that both public
and private methods
being removed from a project is more likely to occur after a major release than after. Table~\ref{tab:ratio} also shows significance in the changes to private
classes. We see that private classes are added more (24\%) before a major release and removed more after (44\%) the release. 
All results in Table~\ref{tab:ratio} could be used as identified trends of major software releases, while we have just highlighted the larger ratios
which meet our threshold criteria in this paper.
